A Constructive Ordered Predicate Logic for Category Theory

### 2022-10-14 paper/paper.tex
appendix work

In Figure~\ref{fig:subst} we give full definitions for well-formedness
of contexts and substitutions.

@@

      \inferrule*[right=TyCtxForm]
        {}
        {\Gamma \isadtctx}

    \inferrule*[right=SetCtxForm]
    {\Gamma \isadtctx}
    {\Gamma \vdash \Xi \boundary}

    \inferrule*[right=TransCtxForm]
    {\Gamma \isadtctx}
     {\Gamma \vdash \Phi \isavectx}
    \inferrule
    {\Delta \isadtctx \and \Gamma \isadtctx}
    {\gamma : \Delta \to \Gamma}

    \inferrule*[right=Boundary Map formation]
    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
    {\Gamma \vdash \xi : Z \to \Xi}

    \inferrule*[right=Cat Subst formation]
    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
    {\Gamma \vdash \phi : \Psi \to \Phi}
  \end{mathpar}
  \begin{mathpar}
    \inferrule{}{\cdot \isadtctx}

    \inferrule
    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
    {\Gamma , X : A \isadtctx}\\

    \inferrule
    {X : A \in \Gamma}
    {\Gamma \vdash X : A}

    \inferrule
    {}
    {\alpha: \cat C \vdash \alpha: \cat C}

    \inferrule
    {\Gamma \vdash \cat C : \isaCat}
    {\Gamma \vdash \alpha: \cat C \isavectx}

    \inferrule
    {\Gamma \vdash \Phi \isavectx\and
    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}

    \inferrule
    {}
    {\Gamma \pipe \alpha:\cat C, x : R, \beta: \cat D \vdash x : R}

    % horizontal identity
    \inferrule
    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}

    \inferrule
    {\Psi = \Psi_1 \jnctx \Psi_2\and
      \Gamma\pipe\Psi_1 \vdash \phi : \Phi\and
      \Gamma \pipe \Psi_2 \vdash t : R[d^+\phi;b/\beta] \and
      \Gamma \vdash \phi_2 : \Psi_2 \to \Phi_2
      }
    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
  \end{mathpar}

  \caption{Substitution}
\end{figure}

\begin{figure}
  \begin{align*}
    \underline{\alpha:C} &= \alpha:C\\
    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
  \end{align*}
  \begin{align*}
    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
  \end{align*}
  \begin{align*}
    d^{\pm}(a/\alpha) &= a/\alpha\\
    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
  \end{align*}
  \caption{Operations on Contexts}
\end{figure}

In Figure~\ref{fig:admissible}, we describe the admissible typing
rules for the actions of substitution.

\begin{definition}[Substitution Actions]
  For any $\Gamma \pipe \alpha: \cat C \vdash a : \cat D$ and $\Gamma
  \pipe \beta:\cat D \vdash b \cat E$, we define $b[a/\beta]$ by
  recursion on $b$:

@@

@@

%% \begin{definition}[Substitution of Objects in Sets]
%%   3
%% \end{definition}
%% \begin{definition}[Substitution of Elements in Elements]
%%   3
%% \end{definition}

\begin{figure}
  \begin{mathpar}
    \inferrule
    {}
    {\cdot : \Gamma \to \cdot}

    \inferrule
    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
    {\gamma,M/X : \Delta \to \Gamma,X:A}

>>>>

In Figure~\ref{fig:ctx} we include the formation rules and definitions
of the three kinds of contexts that are used in \vett{}.
%
In Figure~\ref{fig:substitutions} we give definitions for
well-typedness of the corresponding three kinds of substitutions.

\begin{figure}
  \begin{mathpar}
    \inferrule*[right=TyCtxForm]
    {}
    {\Gamma \isadtctx}

    \inferrule*[right=MtTyCtx]{}{\cdot \isadtctx}

    \inferrule*[right=TyCtxExt]
    {\Gamma \isadtctx \and \Gamma \vdash A \isaTy}
    {\Gamma , X : A \isadtctx}

    \inferrule*[right=BoundaryForm]
    {\Gamma \isadtctx}
    {\Gamma \vdash \Xi \boundary}

    \inferrule*[right=BoundarySingle]
    {\Gamma \vdash \cat C \isaCat}
    {\Gamma \vdash \alpha: \cat C \boundary}

    \inferrule*[right=BoundaryDbl]
    {\Gamma \vdash \cat C \isaCat \and \Gamma\vdash\cat D \isaCat}
    {\Gamma \vdash \alpha: \cat C; \beta: \cat D \isaboundary}

    \inferrule*[right=SetCtxForm]
    {\Gamma \isadtctx}
    {\Gamma \vdash \Phi \isavectx}

    \inferrule*[right=SetCtxMt]
    {\Gamma \vdash \cat C \isaCat}
    {\Gamma \vdash \alpha: \cat C \isavectx}

    \inferrule*[right=SetCtxExt]
    {\Gamma \vdash \Phi \isavectx\and
    \Gamma \pipe d^+\Phi; \beta:\cat D \vdash R \isaSet}
    {\Gamma \vdash \Phi, x:R, \beta:\cat D \isavectx}
  \end{mathpar}
  \caption{Contexts}
  \label{fig:ctx}
\end{figure}

@@

    \inferrule*[right=TermSubstFormation]
    {\Delta \isadtctx \and \Gamma \isadtctx}
    {\Delta \vdash \gamma :: \Gamma}

    \inferrule*[right=TermSubstMt]
    {}
    {\cdot : \Gamma \to \cdot}

    \inferrule*[right=TermSubstExt]
    {\gamma : \Delta \to \Gamma \and \Delta \vdash M : A[\gamma]}
    {\gamma,M/X : \Delta \to \Gamma,X:A}

    \inferrule*[right=BoundarySubstFormation]
    {\Gamma \vdash Z \boundary \and \Gamma \vdash \Xi \boundary}
    {\Gamma\pipe Z \vdash \xi :: \Xi}

    \inferrule*[right=BoundarySubstSingle]
    {\Gamma\pipe \alpha:\cat C \vdash b : \cat D}
    {\Gamma \pipe \alpha:\cat C \vdash b/\beta :: \beta:\cat D}

    \inferrule*[right=BoundarySubstDbl]
    {\Gamma\pipe d^-\Xi \vdash a : \cat C\and
     \Gamma\pipe d^+\Xi \vdash b : \cat D}
    {\Gamma \pipe \Xi \vdash a/\alpha;b/\beta :: (\alpha : \cat C; \beta:\cat D)}

    \inferrule*[right=ElementSubstFormation]
    {\Gamma \vdash \Psi \isavectx \and \Gamma \vdash \Phi \isavectx}
    {\Gamma\pipe \Psi \vdash \phi :: \Phi}

    % horizontal identity
    \inferrule*[right=ElementSubstMt]
    {\Gamma\pipe \beta:\cat C \vdash a : \cat C}
    {\Gamma \pipe \beta:\cat C \vdash a/\alpha : (\alpha : \cat C)}

    \inferrule*[right=ElementSubstExt]
    {\Gamma\pipe\Psi \vdash \phi :: \Phi \and
      d^+\Psi = d^-\Psi'
     \Gamma\pipe\Psi' \vdash t : R[d^+\phi;b/\beta]\and
     \Gamma\pipe d^+\Psi'\vdash b : \cat D}
    {\Gamma\pipe\Psi \vdash \phi,t/x,b/\beta :: (\Phi,x:R,\beta:\cat D)}
  \end{mathpar}
  \caption{Substitution}
  \label{fig:substitions}
\end{figure}

These definitions involve several operations $d^\pm\Phi$, $\underline
\Phi$ and $\Phi\jnctx \Psi$ on contexts (and their functorial lift to
substitutions) that we now define.

\begin{definition}
  We define operations $d^\pm$ that project out the covariant and
  contravariant boundary of a set context.
  This can be typed with the admissible rule
  \begin{mathpar}
    \inferrule*[right=(*)]
               {\Gamma \vdash \Phi \vectx}
               {\Gamma \vdash d^\pm\Phi \boundary}
  \end{mathpar}
  and is defined as
  \begin{align*}
    d^{\pm}(\alpha:\cat C) &= \alpha : \cat C\\
    d^{-}(\Phi,x:R,\Psi) &= d^-\Phi\\
    d^{+}(\Phi,x:R,\Psi) &= d^-\Psi
  \end{align*}
  This operation extends to the substitutions with admissible rule
  \begin{mathpar}
    \inferrule*[right=(*)]
    {\Gamma\pipe\Psi \vdash \phi :: \Phi}
    {\Gamma\pipe d^\pm \Psi \vdash d^\pm\phi :: d^\pm \Phi}
  \end{mathpar}
  defined as
  \begin{align*}
    d^{\pm}(a/\alpha) &= a/\alpha\\
    d^{-}(\phi,t/x,\psi) &= d^-\phi\\
    d^{+}(\phi,t/x,\psi) &= d^+\psi\\
  \end{align*}
\end{definition}

\begin{definition}
  We define the operation of restricting a set context to both sides
  of its boundary with admissible typing
  \begin{mathpar}
    \inferrule*[right=(*)]
    {\Gamma \vdash \Phi \isavectx}
    {\Gamma \vdash \underline\Phi \boundary}
  \end{mathpar}
  and definition
  \begin{align*}
    \underline{\alpha:C} &= \alpha:C\\
    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
  \end{align*}
  The extension to substitutions has admissible typing
  \begin{mathpar}
    \inferrule*[right=(*)]
    {\Gamma \pipe \Psi \vdash \phi :: \Phi}
    {\Gamma\pipe\underline \vdash \Phi \boundary}
  \end{mathpar}
  and definition
  \begin{align*}
    \underline{\alpha:C} &= \alpha:C\\
    \underline{\Phi,x:R,\Psi} &= d^-\Phi; d^+{\Psi}
  \end{align*}
\end{definition}

% RETURN
Finally, we define the operation of ``horizontal composition'' of set
contexts $\Phi \jnctx \Psi$ and its functorial lift $\phi \jnctx
\psi$. Note that the horizontal composition is only well-defined when $$

Next, we define the actions of substitutions on terms. We elide the
obvious action of term substitutions $\gamma$ and include only the
more unusual substructural substitutions.
\begin{definition}[Substitution Actions]
  For any $\Gamma \pipe \alpha: \cat C \vdash a : \cat D$ and $\Gamma
  \pipe \beta:\cat D \vdash b \cat E$, we define $b[a/\beta]$ by
  recursion on $b$:

@@

\begin{lemma}[Properties of Substitution]
  Substitution is associative and unital?
\end{lemma}

@@

\begin{figure}
  \begin{mathpar}
